What public transit API logs tell us about travel flows

### 2016-01-15 paper.tex
Not our data and small abstract change

\maketitle
\begin{abstract}
%Public transit schedules are made available on the Web in various ways, such as in a GTFS file, a route planning API, or through Linked Connections.
%This webservice provides an answer to a route planning question with parameters such as ``departuretime'', ``from'' and ``to''.
% Context & % Need
In the field of urban planning, researchers need an indication of how people move between cities. 
Yet, getting statistics of travel flows within public transit systems has proven to be troublesome.
% Task
In order to get an indication of travel flows between cities in Belgium,
we analyzed the query logs of the iRail API, a highly expressive route planning API for the Belgian railways.
We were able to study $\sim$100k to 500k requests for each month between October 2012 and November 2015, which is between $0.56\%$ and $1.66\%$ of the amount of monthly passengers.
% Object & Findings
Using data visualizations, we illustrate the commuting patterns in Belgium and confirm that Brussels, the capital, acts as a central hub. 
The Flemish region appears to be polycentric, while in the Walloon region, everything converges on Brussels.
The findings correspond to the real travel demand, according to experts of the passenger federation Trein Tram Bus.
% Conclusion
We conclude that query logs of route planners are of high importance in getting an indication of travel flows.
% Perspectives
However, better travel intentions would be acquirable using dedicated HTTP POST requests.
%%Redundant according to RV:
%Highly expressive transport data publishing methods such as route planning API exist, as well as data dumps or data fragments.
%In order to be able to gather meaningful logs in all cases, we suggest using a separate POST request containing the entire query.

@@

%Visualizations are crucial for understanding the patterns in data and representing transit flows.
Visualizations are crucial for making data analysis tangible~\cite{thomas2006visual}, so that results can be communicated and debated~\cite{robinson2008collaborative}. 
This is key for our research, since the questions presented are exploratory in nature~\cite{kraak2008exploratory} and can be addressed in many ways. 
Movement data has received substantial attention from the visualization community and a range of techniques has been developed to support flow analysis~\cite{andrienko2012visual}. 
Visualizations of flows can be broadly categorized into two groups. 
Representations that display the complete trajectory and Origin Destination (OD) type representations that take only the start and end locations into account.
Since our data solely consists of OD information, we narrow our description to the latter group of visualization techniques.

@@

Using this data, everyone can build their own historic query log dataset.
The server costs for iRail to host this dataset stay limited.
Furthermore, this way of publishing data enables research for real-time predictions of travel intentions.
Privacy preservation is paramount~\cite{silvestri} when dealing with query logs and, therefore, IP addresses have been removed to protect the privacy of our users. 
Additionally, the integrity of this dataset, without IP addresses, has been vetted by the Belgian Privacy Commission and permission has been granted for these query logs to be published as open data.

@@

There has been an increase in the number of queries made between 2012 and 2015. 
\cref{fig:querylogs} provides a break down of the numbers on a monthly basis. 
A distinctive increase in the number of queries can be observed in April 2015 when the official route planning service provided by of the Belgian rail company was discontinued\footnote{\url{https://hello.irail.be/2015/04/22/april-updates/}}, resulting in widespread adoption of alternative software applications built on top of the iRail API.
Calculated with indicators provided by the Flemish regional government of Belgium\footnote{\url{http://www4.vlaanderen.be/sites/svr/Cijfers/Exceltabellen/mobiliteit/vervoersprestaties/personenvervoer/MOBIOPEN006.xls}}, we infer a monthly average of 19.3, 19.4, 19,6 and 19.7 million passengers in 2012, 2013, 2014 and 2015 respectively. 
Correspondingly, the average number of iRail queries per month amounts to 0.11, 0.17, 0.15 and 0.33 million respectively.
Assuming that each request reciprocates with an intention to travel by rail, our data captures a respective $0.56\%$, 0.88\%, 0.77\% and 1.66\% of the actual journeys that have occurred.

%After analysis, it appeared 10\% of the request where performed by Google bots.

>>>>

\maketitle
\begin{abstract}
%Public transit schedules are made available on the Web in various ways, such as in a GTFS file, a route planning API, or through Linked Connections.
%This webservice provides an answer to a route planning question with parameters such as ``departuretime'', ``from'' and ``to''.
% Context & % Need
In the field of smart cities, researchers need an indication of how people move in and between cities.
Yet, getting statistics of travel flows within public transit systems has proven to be troublesome.
% Task
In order to get an indication of public transit travel flows in Belgium,
we analyzed the query logs of the iRail API, a highly expressive route planning API for the Belgian railways.
We were able to study $\sim$100k to 500k requests for each month between October 2012 and November 2015, which is between $0.56\%$ and $1.66\%$ of the amount of monthly passengers.
% Object & Findings
Using data visualizations, we illustrate the commuting patterns in Belgium and confirm that Brussels, the capital, acts as a central hub. 
The Flemish region appears to be polycentric, while in the Walloon region, everything converges on Brussels.
The findings correspond to the real travel demand, according to experts of the passenger federation Trein Tram Bus.
% Conclusion
We conclude that query logs of route planners are of high importance in getting an indication of travel flows.
% Perspectives
However, better travel intentions would be acquirable using dedicated HTTP POST requests.
%%Redundant according to RV:
%Highly expressive transport data publishing methods such as route planning API exist, as well as data dumps or data fragments.
%In order to be able to gather meaningful logs in all cases, we suggest using a separate POST request containing the entire query.

@@

%Visualizations are crucial for understanding the patterns in data and representing transit flows.
Visualizations are crucial for making data analysis tangible~\cite{thomas2006visual}, so that results can be communicated and debated~\cite{robinson2008collaborative}. 
This is key for our research, since the questions presented are exploratory in nature~\cite{kraak2008exploratory} and can be addressed in many ways. 
Movement data has received substantial attention from the visualization community and a range of techniques has been developed to support flow analysis~\cite{andrienko2012visual}. 
Visualizations of flows can be broadly categorized into two groups. 
Representations that display the complete trajectory and Origin Destination (OD) type representations that take only the start and end locations into account.
Since iRail's query log data solely consists of OD information, we narrow our description to the latter group of visualization techniques.

@@

Using this data, everyone can build their own historic query log dataset.
The server costs for iRail to host this dataset stay limited.
Furthermore, this way of publishing data enables research for real-time predictions of travel intentions.
Privacy preservation is paramount~\cite{silvestri} when dealing with query logs and, therefore, IP addresses have been removed to protect the privacy of iRail's users. 
Additionally, the integrity of this dataset, without IP addresses, has been vetted by the Belgian Privacy Commission and permission has been granted for these query logs to be published as open data.

@@

There has been an increase in the number of queries made between 2012 and 2015. 
\cref{fig:querylogs} provides a break down of the numbers on a monthly basis. 
A distinctive increase in the number of queries can be observed in April 2015 when the official route planning service provided by of the Belgian rail company was discontinued\footnote{\url{https://hello.irail.be/2015/04/22/april-updates/}}, resulting in widespread adoption of alternative software applications built on top of the iRail API.
Calculated with indicators provided by the Flemish regional government of Belgium\footnote{\url{http://www4.vlaanderen.be/sites/svr/Cijfers/Exceltabellen/mobiliteit/vervoersprestaties/personenvervoer/MOBIOPEN006.xls}}, we infer a monthly average of 19.3, 19.4, 19,6 and 19.7 million passengers in 2012, 2013, 2014 and 2015 respectively. 
Correspondingly, the average number of iRail queries per month amounts to 0.11, 0.17, 0.15 and 0.33 million respectively.
Assuming that each request reciprocates with an intention to travel by rail, iRail's query log data captures a respective $0.56\%$, 0.88\%, 0.77\% and 1.66\% of the actual journeys that have occurred.
%After analysis, it appeared 10\% of the request where performed by Google bots.
